---
title: "linregcode"
author: "Ryley"
date: "2024-04-23"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Graphing Predictor Variables Against the Response Variable

```{r}
library(ggplot2)
ggplot(healthcareins, aes(x = age, y = charges, color = smoker)) +
 geom_point() +
 labs(title = "Charges vs. Age by Smoker Status")

ggsave("C:\\Users\\ryley\\OneDrive\\Desktop\\GitHub\\Heathcare Insurance Predictions\\Figures\\ChargesVAgeBySmoker.png",
  width = 8,
  height = 6,
)
```

```{r}
ggplot(healthcareins, aes(x = children, y = charges, color=smoker)) +
 geom_point() +
 labs(title = "Charges vs. children by Smoker Status")

```

```{r}
ggplot(healthcareins, aes(x = region, y = charges)) +
 geom_boxplot() +
 labs(title = "Charges by Region")

```

```{r}
ggplot(healthcareins, aes(x = age, y = charges, color = sex)) +
 geom_point(alpha = 0.6) +
 facet_wrap(~smoker) +
 labs(title = "Charges vs. Age, Faceted by Smoker Status and Colored by Sex")

```

```{r}
ggplot(healthcareins, aes(x = bmi, y = charges, color = smoker)) +
 geom_point(alpha = 0.6) +
 geom_smooth(method = "lm") +
 labs(title = "Interaction of BMI and Smoker Status on Charges")

```

```{r}
ggplot(healthcareins, aes(x = charges, fill = smoker)) +
 geom_density(alpha = 0.5) +
 facet_grid(smoker ~ .) +
 labs(title = "Density of Charges by Smoker Status")

```

Full Model

```{r}
fullmodel <- lm(charges ~ age + sex + bmi + children + smoker + region, data = healthcareins)
summary(fullmodel)

```

Interactions:

```{r}
# Base model with Age and BMI
simple_model_age_bmi <- lm(charges ~ age + bmi, data = healthcareins)
# Base model with Smoker and Age
simple_model_smoker_age <- lm(charges ~ smoker + age, data = healthcareins)
# Base model with Smoker and BMI
simple_model_smoker_bmi <- lm(charges ~ smoker + bmi, data = healthcareins)
```

```{r}
# Base model with Age and BMI
interaction_model_age_bmi <- lm(charges ~ age + bmi + age * bmi, data = healthcareins)
# Base model with Smoker and Age
interaction_model_smoker_age <- lm(charges ~ smoker + age + smoker * age, data = healthcareins)
# Base model with Smoker and BMI
interaction_model_smoker_bmi <- lm(charges ~ smoker+ bmi+ smoker * bmi, data = healthcareins)
```

```{r}
# ANOVA for Age and BMI
anova_age_bmi <- anova(simple_model_age_bmi, interaction_model_age_bmi)
anova_age_bmi

```

```{r}
# ANOVA for Smoker and Age
anova_smoker_age <- anova(simple_model_smoker_age, interaction_model_smoker_age)
anova_smoker_age
```

```{r}
# ANOVA for Smoker and BMI
anova_smoker_bmi <- anova(simple_model_smoker_bmi, interaction_model_smoker_bmi)
anova_smoker_bmi
```

```{r}
simple_model_age_sex_bmi <- lm(charges ~ age + sex + bmi, data = healthcareins)
interaction_model_age_sex_bmi <- lm(charges ~ sex + bmi+ age + age * bmi+ bmi * sex+ age* sex + age * bmi * sex, data = healthcareins)
# Compare the two models using ANOVA
model_comparison <- anova(simple_model_age_sex_bmi, interaction_model_age_sex_bmi)
model_comparison

```

```{r}
simple_model_age_smoker_bmi <- lm(charges ~ smoker + bmi+ age, data = healthcareins)
interaction_model_age_smoker_bmi <- lm(charges ~ smoker + bmi+ age + age * bmi+ bmi * smoker+ age* smoker + age * bmi * smoker, data = healthcareins)

model_comparison2 <- anova(simple_model_age_smoker_bmi, interaction_model_age_smoker_bmi)
model_comparison2
```

```{r}
summary(interaction_model_age_smoker_bmi)
```

```{r}
interactions=lm(charges~.^2,data=healthcareins[,-1])
step(interactions, direction = "both")
```

```{r}
interaction_model_full=lm(charges ~ sex + bmi + children + smoker + region + sex*children + bmi*smoker + bmi*region, healthcareins)
summary(interaction_model_full)
```




backwards

```{r}
step(fullmodel, direction = "backward")
```

forwards

```{r}
ins_mod_start = lm(charges ~ 1, data = healthcareins)
ins_mod_forw_aic = step(
  ins_mod_start, 
  scope = charges ~ age + sex + bmi + children + smoker + region, 
  direction = "forward")
```

stepwise

```{r}
step(fullmodel, direction = "both")
```

```{r}
AIC_BIC_model=lm(charges ~ age + bmi + children + smoker + region, healthcareins)
summary(AIC_BIC_model)
```



Quadratic

```{r}
quadratic = lm(charges~age+bmi+children+sex+smoker+region+I(age^2)+I(bmi^2), healthcareins)
summary(quadratic)

```

```{r}
#install.packages("leaps")
install.packages("olsrr")
```

best subset selection:

```{r}
library(leaps)
all=regsubsets(charges~.,healthcareins,nvmax=6)
```

```{r}
plot(summary(all)$rsq ~ seq(6), type = "b", col = "red")


```

```{r}
plot(summary(all)$adj ~ seq(6), type = "b", col = "red")

```

```{r}
plot(summary(all)$cp ~ seq(6), type = "b", col = "red")

```

```{r}
plot(summary(all)$bic ~ seq(6), type = "b", col = "red")

```

```{r}
library(olsrr)
FitAll <- lm(charges ~ ., healthcareins)
bestAIC <- ols_step_best_subset(FitAll)
values=bestAIC$metrics$aic
plot(values ~ seq(6), type = "b", col = "red")
```

All of the best subset selection charts advise to choose 3 predictor variables for the best model

```{r}
summary(all)
```

From this summary we can see the variables are smoker, age, and bmi

```{r}
bestsubsetmodel=lm(charges ~ smoker + age+ bmi, healthcareins)
summary(bestsubsetmodel)
```

cross validation

```{r}

library(DAAG)
```

```{r}
#CVlm(healthcareins,charges ~  age+ bmi + smoker + age* bmi *smoker, m=3)
```

ms=24356280

```{r}
#CVlm(healthcareins,charges ~  sex + bmi + children + smoker + region + sex*children + bmi*smoker + bmi*region, m=3)
```

ms=37159125

```{r}
#CVlm(healthcareins, charges ~  age + bmi + children + smoker + region, m=3)
```

ms=37217608

```{r}
#CVlm(healthcareins,charges ~  age+ bmi + children + smoker + region + I(age^2) + I(bmi^2), m=3)
```

ms = 36703796

```{r}
#CVlm(healthcareins,charges ~  age+ bmi + smoker, m=3)
```

ms=37380135

Final Model

```{r}
interaction_model_age_smoker_bmi <- lm(charges ~ smoker + bmi+ age + age * bmi+ bmi * smoker+ age* smoker + age * bmi * smoker, data = healthcareins)
summary(interaction_model_age_smoker_bmi)
```

Intercept: The predicted average healthcare Insurance Charge for non-smokers is -1.483e+03

slope for smokeryes: The predicted average difference between the healthcare insurance charges for smokers and nonsmokers is -1.242e+04

slope for bmi: For every 1 unit increase in bmi, healthcare insurance charges decrease by -1.943e+01

slope for age: For every 1 unit increase in age, healthcare insurance charges increase by 2.458e+02

interaction for bmi and age: The difference in slopes between bmi and age has an estimate of approximately 6.835e-01

interaction for bmi and smokeryes: The difference in slopes between being a smoker and bmi has an estimate of approximately 1.184e+03

interaction for age and smokeryes: The difference in slopes between being a smoker and age has an estimate of approximately -2.068e+02

interaction for age, smokeryes, and bmi: The difference in slopes between being a smoker, age, and bmi has an estimate of approximately 6.624e+00

Rsq = 83.68% of the variation in healthcare insurance charges can be explained by the final model

$H_{0}: \beta_{1}=\beta_{2}=\beta_{3}=\beta_{4}=\beta_{5}=\beta_{6}=\beta_{7}=\beta_{8}=0$ The model is not useful in predicting healthcare insurance charges $H_{A}:$ at least one $\beta_{i} \neq 0$ the model is useful in predicting healthcare insurance charges

p-value is less than 2.2e-16 and less than $\alpha = 0.01$ so we reject the $H_{0}$ and conclude that the model is useful in predicting the healthcare insurance charges



